\documentclass[11pt]{article}
\input{_preamble}

\title{Daily Exercise}
\author{Dhruman Gupta}
\date{30th March}

\begin{document}

\input{_exercise-header}
Let $L=\{\langle M\rangle \mid |L(M)|=72\}$, i.e., the language of encodings of Turing machines whose accepted language contains exactly 72 strings. Prove that $L$ is neither recursively enumerable nor co-recursively enumerable.

\textbf{Answer:}

First, recall that
\[
E_{TM} = \{\langle M \rangle \mid L(M) = \emptyset\}
\]
is not recursively enumerable.

To see this, reduce from $\overline{A_{TM}}$. Given $\langle M,w \rangle$, construct a machine $N$ as follows: on any input $x$,
\begin{itemize}
    \item Simulate $M$ on $w$.
    \item If $M$ accepts $w$, then accept $x$.
    \item Otherwise, run forever.
\end{itemize}
Then
\[
\langle M,w \rangle \in \overline{A_{TM}} \iff L(N)=\emptyset.
\]
Hence $\overline{A_{TM}} \leq_m E_{TM}$, so $E_{TM} \notin RE$.

Now we prove that
\[
L = \{\langle M \rangle \mid |L(M)| = 72\}
\]
is neither recursively enumerable nor co-recursively enumerable.

First, show that $L \notin RE$.

Take any machine $M$. Construct a machine $N$ as follows. On input $x$:
\begin{itemize}
    \item If $x \in \{0,0^2,\dots,0^{72}\}$, accept.
    \item Otherwise, dovetail $M$ on all strings $w_1,w_2,\dots$
    \item If some $w_i$ is accepted by $M$, then accept if $x=1$.
    \item Otherwise run forever.
\end{itemize}

If $L(M)=\emptyset$, then the dovetailing never finds an accepted string, so $N$ accepts exactly
\[
\{0,0^2,\dots,0^{72}\},
\]
and therefore $|L(N)|=72$.

If $L(M)\neq \emptyset$, then eventually the dovetailing finds some accepted string, and then $N$ also accepts $1$. Hence in this case $|L(N)|=73$.

So
\[
\langle M \rangle \in E_{TM} \iff \langle N \rangle \in L.
\]
Thus
\[
E_{TM} \leq_m L.
\]
Since $E_{TM}\notin RE$, it follows that $L\notin RE$.

Now show that $L \notin coRE$. Equivalently, show that $\overline{L}\notin RE$.

Again take any machine $M$, and construct a machine $N$ as follows. On input $x$:
\begin{itemize}
    \item If $x \in \{0,0^2,\dots,0^{71}\}$, accept.
    \item Otherwise, dovetail $M$ on all strings $w_1,w_2,\dots$
    \item If some $w_i$ is accepted by $M$, then accept if $x=1$.
    \item Otherwise run forever.
\end{itemize}

If $L(M)=\emptyset$, then $N$ accepts exactly
\[
\{0,0^2,\dots,0^{71}\},
\]
so $|L(N)|=71$, and therefore $\langle N\rangle \in \overline{L}$.

If $L(M)\neq \emptyset$, then eventually $N$ also accepts $1$, so $|L(N)|=72$, and hence $\langle N\rangle \in L$.

Therefore
\[
\langle M \rangle \in E_{TM} \iff \langle N \rangle \in \overline{L}.
\]
So
\[
E_{TM} \leq_m \overline{L}.
\]
Since $E_{TM}\notin RE$, we get $\overline{L}\notin RE$. Hence $L\notin coRE$.

Therefore $L$ is neither recursively enumerable nor co-recursively enumerable.

\end{document}
